---
title: "Subgroup calibration - Case 000"
author: "Huiding Chen"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: "hide"
---

```{r,include=FALSE}
library(knitr)
library(dplyr)
library(tidyverse)
library(cat)
library(survey)
library(kableExtra)
library(VGAM)
library(distributionsrd)
library(ggplot2)
library(truncnorm)
library(reshape2)
library(paletteer)
library(plotly)
library(stringr)
library(purrr)
library(RVAideMemoire)
library(glmnet)
library(DT)
library(viridis)
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```


# Brief Summary

| Covariate Dependency|Outcome Heterogeneity|Non-prob Heterogeneity|
| --------------------:|:-------------------:|---------------------:|
| No|No|No|


```{r}
spec_color2 <- function(x, alpha = 1, begin = 0, end = 1,
                        direction = 1, option = "D",
                        na_color = "#BBBBBB", scale_from = NULL, palette = NULL
                        ) {
  
  n <- length(palette)
  if(is.null(palette)){
    palette = viridisLite::viridis(256, alpha = 1, begin = 0, end = 1, direction = 1, option = "D")
  }
  if (is.null(scale_from)) {
    x <- round(scales::rescale(x, c(1, n)))
  } else {
    x <- round(scales::rescale(x, to = c(1, n),
                               from = scale_from))
  }
  
  color_code <- palette[x]
  color_code[is.na(color_code)] <- na_color
  return(color_code)
}

# Function to apply row_spec automatically

apply_row_colors <- function(df,names_caption = NULL,
                             n = 1000, size = 1e3,digits = 5) {
  df <- df %>% 
    as.data.frame() %>% 
    select(where(~ !any(is.na(head(.x, -1))))) %>% 
    na.omit %>%
    (`*`)(size) %>%
    round(digits)
  # Apply cell_spec to each cell in the numeric columns
  df_modified <- df %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ cell_spec(
    ., 
    background = spec_color2(., alpha = 1,  
                             option = "D",
                             begin = 0.2,
                             end = 0.6,
                             palette = paletteer_d("ggsci::orange_material",8)), 
    color = "black"
  ))) %>%
  t() %>%
  as.data.frame()
  # Create a kable table with the modified data frame
  table <- df_modified %>%
    kable(format = "html",  align = "c",
          caption = names_caption,
          escape = FALSE) %>%
    kable_styling(full_width = FALSE, c("striped", "condensed"),)
  
  return(table)
}

#names_var <- c("catx1","catx2","catx3")
```

## Intro

```{r}
scenario <- "X5_V12"
case <- "000"
var_method <- c("rails","rails_subgroup")
var_stratified <- c("catx5")
```


```{r}
n = 1000
pre_pop <- read.csv(paste0("pre_pop_",scenario,"_",case,".csv"),header = TRUE)
pre_all <- read.csv(paste0("pre_all_",scenario,"_",case,".csv"))
truey <- as.numeric(pre_pop[,2])[1]
pre_pop <- unlist(pre_pop[,2])[-1]
pre_all <- unlist(pre_all[,2])[-1]
sim_result <- readRDS(paste0(scenario,"_",case,".RData"))

```

## Report Function

```{r}
fun.rep <- function(x = sim_result,n,pre_pop,truey,
                    var_method,var_stratified){
  # x as the input list
  # v as the variable names to be reported
  # n as the number of simulations
  x <- sim_result
  # Define the population
  temp_pop <- c(pre_pop,0)
  
  # Names of the variables
  names_var <- colnames(x[[1]]$unweighted$sub)
  n_var <- length(names_var)
  
  # Names of the Methods
  index_method <- which.max(sapply(x, length))
  names_method <- names(x[[index_method]])[1:(length(x[[index_method]])-3)]
  n_method <- length(names_method)
  names_method_fix <- c("naive","oracle","G-Raking","NPS","GVS",
                        "G-RAILS","S-Raking","S-RAILS")
  
  # Index for population estimate
  names_popest <- names(x[[index_method]]$unweighted$pop)
  n_popest <- length(names_popest)
  
  # Index for subgroup estimate
  names_subest <- rownames(x[[index_method]]$unweighted$sub)
  n_subest <- length(names_subest)
  
  # Count the number of NAs for each method
  na_count <- sapply(names_method, function(method) {
    sum(sapply(x, function(sim) {
      any(is.na(sim[[method]]$sub))
    }))
  })
  
  
  
  # Collect the number of error messages
  error_count <- c()
  error_index <- c()
  for (i in 1:n){
    if(!is.null(x[[i]]$error)){
      error_count <- c(error_count,x[[i]]$error)
      error_index <- c(error_index,i)
    }
  }
  error_count <- as.vector(table(error_count))
  if(length(error_count)==0){
    error_count <- 0
  }
  # # Same as Rails
  # diff_from_rails <- c()
  # for (i in 1:n){
  #   if(!is.null(x[[i]]$mix_RR$diff_from_rails)){
  #     diff_from_rails <- c(diff_from_rails,x[[i]]$mix_RR$diff_from_rails)
  #   }
  # }
  # diff_from_rails <- sum(diff_from_rails)
  
  # Population Prevalence
  count_non_na <- numeric(length = n_method)
  sum_pop <- matrix(0,ncol = n_popest,nrow = n_method)
  est_pop <- matrix(0,nrow = n,ncol = n_method)
  sd_pop <- matrix(0,nrow = n,ncol = n_method)
  colnames(sum_pop) <- names_popest
  colnames(est_pop) <- names_method_fix
  colnames(sd_pop) <- names_method_fix
  rownames(sum_pop) <- names_method_fix
  for(i in setdiff(1:n,error_index)){
    for(j in 1:n_method){
      temp <- x[[i]][[names_method[j]]]$pop
      count_non_na[j] <- count_non_na[j] + all(!is.na(temp[1]))
      sum_pop[j,] <- sum_pop[j,] + ifelse((is.na(temp)),0,temp)
      est_pop[i,j] <-  ifelse(is.na(temp[1]),NA,temp[1])
      sd_pop[i,j] <-  ifelse(is.na(temp[2]),NA,sqrt(temp[2]))
    }
  }
  out_mean <-sum_pop/count_non_na
  out_mean <- cbind(out_mean, "Ratio of NAs" = 1 - count_non_na/n)
  out_mean <- as.data.frame(out_mean)
  empsd <-      apply(est_pop,2,sd,na.rm = TRUE)
  est_median <- apply(est_pop,2,median,na.rm = TRUE)
  exp_sd <-     apply(sd_pop,2,mean,na.rm = TRUE)
  sd_median <-  apply(sd_pop,2,median,na.rm = TRUE)
  out_mad <-    apply(est_pop,2,mad,na.rm = TRUE)
  
  
  # Bias v.s. true y
  out_mean <- out_mean %>%
    mutate(median = est_median,.after = "mean") %>%
    mutate("Bias(Mean - True)" = mean - truey,.after = bias2) %>%
    mutate("Bias(Median - True)" = median - truey,.after = "Bias(Mean - True)")%>%
    mutate("Relative Bias(Mean)" = (mean - truey)/truey,.after = "Bias(Median - True)") %>%
    mutate("Relative Bias(Median)" = (median - truey)/truey,.after = "Relative Bias(Mean)") %>%
    mutate("Avg SD" = exp_sd,.after = "Relative Bias(Median)") %>%
    mutate("Emp SD" = empsd,.after = "Avg SD") %>%
    mutate("Median SD" = sd_median,.after = "Emp SD") %>%
    mutate("MAD" = out_mad,.after = "Emp SD")
  
  # Oracle Coverage
  out_oracov <- numeric(length = n_method)
  for(i in setdiff(1:n,error_index)){
    for(j in 1:n_method){
      temp <- x[[i]][[names_method[j]]]$pop
      debias_est <- temp[1] - out_mean$`Bias(Mean - True)`[j]
      temp_ci <- debias_est + c(-1,1)*1.96*sqrt(temp[2])
      temp_oracov <- truey >= temp_ci[1] & truey <= temp_ci[2]
      out_oracov[j] <- out_oracov[j] + ifelse(is.na(temp_oracov),0,temp_oracov)
    }
  }
  names(out_oracov) <- names_method_fix
  out_oracov <- out_oracov/count_non_na
  out_mean <- out_mean %>%
    mutate("Oracle Coverage" = out_oracov,.after = TrueCoverage)
  
  # Mean of Bias
  out_list <- vector("list", length(names_subest))
  names(out_list) <- names_subest
  for(k in 1:n_subest){
    out_list[[k]] <- as.data.frame(matrix(0,ncol = n_var ,nrow = n_method))
    colnames(out_list[[k]]) <- names_var
    for(i in setdiff(1:n,error_index)){
      for(j in 1:n_method){
        temp <- unlist(x[[i]][[names_method[j]]]$sub[k,])
        out_list[[k]][j,] <- out_list[[k]][j,] +  ifelse((is.na(temp)),rep(0,n_var),temp)
      }
    }
  }
  for(k in 1:n_subest){
    rownames(out_list[[k]]) <- names_method_fix
    out_list[[k]] <- out_list[[k]]/count_non_na
  }
  out_sub_mean <- out_list$mean[2,]
  out_sub_bias <- sweep(out_list$mean,2, as.numeric(pre_pop), FUN = "-")
  out_sub_bias2 <- (out_sub_bias)^2
  out_sub_cov <- out_list$Coverage
  
  # Elapsed Time
  out_time <- numeric(length = n_method)
  names(out_time) <- names_method_fix
  for(i in setdiff(1:n,error_index)){
    for(j in 1:n_method){
      temp <- x[[i]][[names_method[j]]]$time
      out_time[j] <- out_time[j] + ifelse((is.na(temp)),0,temp)
    }
  }
  out_time <- out_time/count_non_na
  
  # Average Size
  out_size <- vector("numeric", length = n_var)
  #names(out_size) <- names_var
  for(i in setdiff(1:n,error_index)){
    temp <- x[[i]]$avg_size
    out_size <- out_size + temp
  }
  out_size <- out_size/(n-error_count)
  
  # Average Variable Selected
  out_vs_n <- numeric(3)
  for(i in setdiff(1:n,error_index)){
    temp <- x[[i]]$rails$n_stepwise
    out_vs_n <- out_vs_n + ifelse((is.na(temp)),0,temp)
  }
  out_vs_n <- out_vs_n/count_non_na[names_method=="rails"]
  #out_vs_n <- c(total= sum(out_vs_n),out_vs_n)
  
  
  
  # Plot for Total Bias
  bias <- out_mean$`Bias(Mean - True)`
  tt <- data.frame(matrix(0,nrow = n,ncol = n_method))
  colnames(tt) <- names_method_fix
  
  for(i in setdiff(1:n,error_index)){
    for(j in 1:n_method){
      temp <- x[[i]][[names_method[j]]]$pop[1]
      tt[i,j] <-  ifelse((is.na(temp)),NA,temp) - truey
    }
  }
  
  tmat <- data.frame(Method = names_method_fix,MeanBias = bias) %>% na.omit()
  level_order <- tmat$Method
  tmp_all <- tt %>%
    as.data.frame() %>%
    select(where(~ !all(is.na(.)))) %>%
    tidyr::pivot_longer(everything(), names_to = "Method",values_to = "value") %>%
    right_join(tmat) %>%
    mutate(Method = factor(Method, levels = level_order)) %>%
    mutate(Method = droplevels(Method))
  
  # Plot for Subgroup Mean
  out_sub_plot_mean <- vector("list", n_var)
  names(out_sub_plot_mean) <- names_var
  for(k in 1:n_var){
    sub_mean <- out_sub_mean[,k]
    out_sub_plot_mean[[k]] <- as.data.frame(matrix(0,nrow = n ,ncol = n_method))
    colnames(out_sub_plot_mean[[k]]) <- names_method_fix
    for(i in setdiff(1:n,error_index)){
      for(j in 1:n_method){
        temp <- unlist(x[[i]][[names_method[j]]]$sub[2,k])
        out_sub_plot_mean[[k]][i,j] <- ifelse((is.na(temp)),rep(0,n_var),temp)
      }
    }
    tmat <- data.frame(Method = names_method_fix,PopMean = sub_mean) %>% na.omit()
    level_order <- tmat$Method
    tmp <- out_sub_plot_mean[[k]] %>%
      as.data.frame() %>%
      select(where(~ !all(is.na(.)))) %>%
      tidyr::pivot_longer(everything(), names_to = "Method",values_to = "value") %>%
      right_join(tmat) %>%
      mutate(Method = factor(Method, levels = level_order)) %>%
      mutate(Method = droplevels(Method))
    out_sub_plot_mean[[k]] <- tmp
  }
  
  
  
  
  # Mean of Bias
  # Plot for Subgroup Bias
  out_sub_plot_bias <- vector("list", n_var)
  names(out_sub_plot_bias) <- names_var
  for(k in 1:n_var){
    bias <- out_sub_bias[,k]
    out_sub_plot_bias[[k]] <- as.data.frame(matrix(0,nrow = n ,ncol = n_method))
    colnames(out_sub_plot_bias[[k]]) <- names_method_fix
    for(i in setdiff(1:n,error_index)){
      for(j in 1:n_method){
        temp <- unlist(x[[i]][[names_method[j]]]$sub[2,k])
        out_sub_plot_bias[[k]][i,j] <- ifelse((is.na(temp)),rep(0,n_var),temp) - pre_pop[k]
      }
    }
    tmat <- data.frame(Method = names_method_fix,MeanBias = bias) %>% na.omit()
    level_order <- tmat$Method
    tmp <- out_sub_plot_bias[[k]] %>%
      as.data.frame() %>%
      select(where(~ !all(is.na(.)))) %>%
      tidyr::pivot_longer(everything(), names_to = "Method",values_to = "value") %>%
      right_join(tmat) %>%
      mutate(Method = factor(Method, levels = level_order)) %>%
      mutate(Method = droplevels(Method))
    out_sub_plot_bias[[k]] <- tmp
  }
  
  # Histogram for models selected
  out_model <- replicate(1000,
                     vector("list", 1 + length(var_method)),  # expression evaluated 1000×
                     simplify = FALSE)
  diff_from_rails <- rep(1,length = n)
  for(i in setdiff(1:n,error_index)){
    out_model[[i]][[1]] <- x[[i]]$rails$model
    temp_model <- x[[i]]$rails$model
    # Rule out characters if consist of var_startifed for temp_model
    temp_model <- temp_model[!grepl(paste(var_stratified,collapse = "|"),temp_model)]
    if(!is.null(x[[i]]$rails_subgroup$diff_from_rails)){
      diff_from_rails[i] <- 0
    } else {
      exprs <- unlist(lapply(seq_along(var_method[-1]), function(k) {
        nm <- var_method[k+1]
        c(
          sprintf("out_model[[%d]][[2]] <- list(x[[%d]]$%s$model_ori)",
                  i,i, nm),
          sprintf("out_model[[%d]][[3]] <- list(x[[%d]]$%s$model_lifo)",
                  i,i, nm)
        )
      }))
      big_code <- paste(exprs, collapse = "\n")
      eval(parse(text = big_code))
    }
  }
  
  out <- list(out_mean = out_mean,
            out_sub_mean = out_sub_mean,
            out_sub_bias = out_sub_bias,
            out_sub_bias2 = out_sub_bias2,
            out_sub_cov = out_sub_cov,
            out_time = out_time,
            out_size = out_size,
            out_vs_n = out_vs_n,
            out_p_all = tmp_all,
            out_list = out_list,
            out_p_sub_mean = out_sub_plot_mean,
            out_p_sub_bias = out_sub_plot_bias,
            out_level_order = level_order,
            out_error = error_count,
            out_na_count = na_count,
            out_diff_from_rails = diff_from_rails,
            out_model = out_model)
  return(out)
}
```

# Results with Linear Methods

Several methods results are omitted as they don't converge in this setup of simulation. Mostly when the calibration equation contains two-way interactions.

By default, the results are simulated via 1000 replications, except for coverage and relative bias.

```{r}
if(paste0(scenario,"_summary_",case,".RData") %in% list.files(getwd(), recursive = TRUE)){
  out <- readRDS(paste0(scenario,"_summary_",case,".RData"))
} else {
  out <- fun.rep(x = sim_result,
                 n = n,
                 pre_pop = pre_pop,
                 truey = truey,
                 var_method = var_method,
                 var_stratified = var_stratified)
  saveRDS(out, paste0(scenario,"_summary_",case,".RData"))
}
```

## Convergence Information & Error

```{r}
out_error <- out$out_error
out_na_count <- out$out_na_count
out_diff_from_rails <- sum(out$out_diff_from_rails)
kable(out_error,format = "html",caption = "Convergence Information",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"))
kable(out_na_count,format = "html",caption = "Number of NAs",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"))
```


## Population Prevalence

```{r}
out_pop_mean <- as.data.frame(out$out_mean)
out_pop_mean <- out_pop_mean %>% select(!Coverage) %>% t()
apply_row_colors(out_pop_mean,names_caption = "Population Prevalence",size = 1)
```

```{r}
out_pop_mean <- out_pop_mean %>%
  as.data.frame() 
out_pop_mean <- out_pop_mean[c("Relative Bias(Mean)","Relative Bias(Median)","Avg SD","Emp SD","Median SD","MAD","TrueCoverage","Oracle Coverage","Ratio of NAs"),]
out_pop_mean[1,] <- out_pop_mean[1,]*100
out_pop_mean[2,] <- out_pop_mean[2,]*1e2
out_pop_mean[3,] <- out_pop_mean[3,]*1e2
out_pop_mean[4,] <- out_pop_mean[4,]*1e2
out_pop_mean[5,] <- out_pop_mean[5,]*100
out_pop_mean[6,] <- out_pop_mean[6,]*100

apply_row_colors(out_pop_mean,names_caption = "Population Prevalence",size = 1)
```


## Subgroup Bias

```{r}
out_sub_bias <- t(out$out_sub_bias)
out_sub_bias <- rbind(out_sub_bias,Total = colSums(out_sub_bias))
apply_row_colors(out_sub_bias,names_caption = "Subgroup Bias",size = 1e2)
```

## Subgroup Bias Abosulte

```{r}
out_sub_absbias <- abs(t(out$out_sub_bias))
out_sub_absbias <- rbind(out_sub_absbias,Total = colSums(out_sub_absbias))
apply_row_colors(out_sub_absbias,names_caption = "Subgroup Bias Absolute",size = 1e2)
```



## Subgroup Bias Squared

```{r}
out_sub_bias2 <- t(out$out_sub_bias2)
out_sub_bias2 <- rbind(out_sub_bias2,Total = colSums(out_sub_bias2))
apply_row_colors(out_sub_bias2,names_caption = "Subgroup Bias Squared",size = 1e2)
```

## Subgroup Coverage

```{r}
out_sub_cov <- t(out$out_sub_cov)
out_sub_cov <- rbind(out_sub_cov,Total = colMeans(out_sub_cov))
apply_row_colors(out_sub_cov,names_caption = "Subgroup Coverage",size = 1e2)
```

## Subgroup Relative Bias

```{r}
out_sub_relbias <- sweep(out_sub_bias,2, as.numeric(pre_pop), FUN = "/")
out_sub_relbias <- t(out_sub_relbias)
out_sub_relbias <- rbind(out_sub_relbias,Total = colSums(out_sub_relbias))
apply_row_colors(out_sub_relbias,names_caption = "Subgroup Relative Bias",size = 1e2)
```


## Subgroup Summary
```{r}
library(dplyr)
library(tidyr)
out_list <- out$out_list
out_list$Coverage <- out_list$Coverage * 100
out_list$variance <- out_list$variance * 1e6
out_list$size <- out_list$size / 1e5

long_df <- bind_rows(
  lapply(names(out_list), function(m) {
    as.data.frame(out_list[[m]]) |>
      mutate(Method = rownames(out_list[[m]]),
             Metric = m)
  })
) |>
  pivot_longer(
    cols = starts_with("V"),
    names_to = "V",
    values_to = "Value"
  )
tables_by_V <- split(long_df, long_df$V)
tables_by_V_wide <- lapply(tables_by_V, function(df) {
  df %>%
    select(-V) %>%  # we already know the V (table name)
    pivot_wider(names_from = Metric, values_from = Value)
})

# Example: view the wide table for V1

tab_V1 <- tables_by_V_wide$V1 %>% arrange(Method)
tab_V2 <- tables_by_V_wide$V2 %>% arrange(Method)
tab_V3 <- tables_by_V_wide$V3 %>% arrange(Method)
tab_V4 <- tables_by_V_wide$V4 %>% arrange(Method)

kable(tab_V1 ,format = "html",caption = "Subgroup Summary for V1",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
kable(tab_V2 ,format = "html",caption = "Subgroup Summary for V2",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
kable(tab_V3 ,format = "html",caption = "Subgroup Summary for V3",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
kable(tab_V4 ,format = "html",caption = "Subgroup Summary for V4",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
```



## Elapsed Time

```{r}
out_time <- out$out_time
kable(out_time,format = "html",caption = "Elapsed Time (in seconds)",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
```

## Average Size

```{r}
out_size <- out$out_size
kable(out_size,format = "html",caption = "Average Size",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
```

## Average Variable Selected

```{r}
out_vs_n <- out$out_vs_n
out_vs_n_step <- out$out_vs_n_step
out_vs <- cbind(VS = out_vs_n,Step = out_vs_n_step)
kable(out_vs,format = "html",caption = "Average Variables Selected",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"),)
```



## Boxplot

```{r}
tmp <- out$out_p_all
level_order <- out$out_level_order
p <- ggplot(data = tmp, aes(x=Method, y=value,fill=Method))+
    geom_violin(width = 1.2)+
    geom_boxplot(outlier.colour="red",width = 0.5)+
    ggtitle("Boxplot of Bias") +
    xlab("Methods") +
    ylab("Bias") +
    geom_jitter(shape=16, position=position_jitter(0))+
    stat_summary(fun="mean", geom="point", shape=23, size=4,na.rm = TRUE)+
    stat_summary(fun="median", geom="point", shape=1, size=3,na.rm = TRUE)+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
    theme(legend.text = element_text(size=8),
          legend.position="none",
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.ticks.x=element_blank()) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE))+
    scale_x_discrete(limits = level_order) +
    scale_fill_discrete(breaks = level_order,labels = level_order)

p
```

## Subgroup Mean Boxplot

```{r}
tmp <- out$out_p_sub_mean
level_order <- out$out_level_order
tmp <- do.call(rbind, lapply(seq_along(tmp), function(i) {
  cbind(tmp[[i]], Level = i)
}))
tmp <- tmp %>%
  filter(value != 0)

ref_levels <- tmp %>%
  group_by(Level) %>%
  summarise(PopMean  = unique(PopMean), .groups = "drop")

  p <- ggplot(data = tmp, aes(x=Method, y=value,fill=Method))+
    geom_violin()+
    geom_boxplot(outlier.colour="red")+
    geom_hline(aes(yintercept=out$out_mean[1,1], color = "red"), linetype="dashed")+
    geom_hline(data = ref_levels,aes(yintercept=PopMean),color = "blue", linetype="dashed") +
    facet_wrap(~ Level, scales = "free_x", nrow = 1,
               labeller = as_labeller(function(x) paste("Subgroup", x))) +
    ggtitle(paste0("Boxplot of Mean for Subgroup")) +
    xlab("Methods") +
    ylab("Mean") +
    geom_jitter(shape=16, position=position_jitter(0))+
    stat_summary(fun="mean", geom="point", shape=23, size=4,na.rm = TRUE)+
    stat_summary(fun="median", geom="point", shape=1, size=3,na.rm = TRUE)+
    theme(legend.text = element_text(size=8),
          legend.position="none",
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.ticks.x=element_blank()) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE))+
    scale_x_discrete(limits = level_order) +
    scale_fill_discrete(breaks = level_order,labels = level_order) 
  
print(p)




```

## Subgroup Bias Boxplot

```{r}
tmp <- out$out_p_sub_bias
level_order <- out$out_level_order

for(i in 1:length(tmp)){
  p <- ggplot(data = tmp[[i]], aes(x=Method, y=value,fill=Method))+
    geom_violin()+
    geom_boxplot(outlier.colour="red")+
    ggtitle(paste0("Boxplot of Bias for Subgroup ",names(tmp)[i])) +
    xlab("Methods") +
    ylab("Bias") +
    geom_jitter(shape=16, position=position_jitter(0))+
    stat_summary(fun="mean", geom="point", shape=23, size=4,na.rm = TRUE)+
    stat_summary(fun="median", geom="point", shape=1, size=3,na.rm = TRUE)+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
    theme(legend.text = element_text(size=8),
          legend.position="none",
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.ticks.x=element_blank()) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE))+
    scale_x_discrete(limits = level_order) +
    scale_fill_discrete(breaks = level_order,labels = level_order)
  
  print(p)
}



```

## Subgroup HeatMap

```{r}
out_list$Coverage <- out_list$Coverage / 100
coverage_long <- out_list$Coverage |>
  as.data.frame() |>
  rownames_to_column("Method") |>
  pivot_longer(
    cols = starts_with("V"),
    names_to = "Estimand",
    values_to = "Coverage"
  ) |>
  mutate(
    Method = factor(Method, levels = rev(level_order))
  )

ggplot(coverage_long,
       aes(x = Estimand, y = Method, fill = Coverage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.3f", Coverage)), size = 3) +
  scale_fill_gradient2(
    low  = "#EF8A62",   # muted red (under-coverage)
    mid  = "#F7F7F7",   # near-white at nominal
    high = "steelblue",   # muted blue (over / good)
    midpoint = 0.7,
    limits = c(0, 1),
    name = "Coverage"
  ) +
  theme_minimal(base_size = 12)

```



## Histogram of selected variable

### National Rails

```{r}
tmp <- out$out_model
# count
tmp_rails <- c()

for(i in 1:n){
  tmp_rails <- c(tmp_rails,tmp[[i]][[1]])
}
tab <- table(unlist(tmp_rails))

# convert to data.frame
df2 <- as.data.frame(tab, responseName="Freq")


# Ranking
rank_models <- vector("list", n)
for (i in 1:n) {
  rank_models[[i]] <- tmp[[i]][[1]]
}
all_elements <- unique(unlist(rank_models))
element_ranks <- setNames(vector("list", length(all_elements)), all_elements)
# Collect ranks for each element
for (vec in rank_models) {
  for (i in seq_along(vec)) {
    element <- vec[i]
    element_ranks[[element]] <- c(element_ranks[[element]], i)
  }
}
avg_ranks <- sapply(element_ranks, mean)
avg_ranks <- sort(avg_ranks)
avg_rank_df <- data.frame(
  element = names(avg_ranks),
  avg_rank = avg_ranks
)

kable(avg_ranks,format = "html",caption = "Ranking of RAILS",align = "c")%>%
  kable_styling(full_width = FALSE, c("striped", "condensed"))

# Sort the data frame by avg_rank_df
df2$Var1 <- factor(df2$Var1, levels = avg_rank_df$element)
df2 <- df2[order(df2$Var1), ]

# plot
ggplot(df2, aes(x = reorder(Var1, -Freq), y = Freq)) +
  geom_bar(stat = "identity") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
  )+
  scale_fill_brewer(palette="Dark2")+ 
  #scale_x_discrete(limits = avg_rank_df$element) +
  labs(title="Rails by National totals", 
         x="Variables", y = "Frequency") 
```



## Rails by subgroup

```{r}
# count
n_stratified <- length(out$out_size)
tmp_rails_subgroup_ori <-  vector("list",  n_stratified)
tmp_rails_subgroup_final <- vector("list",  n_stratified)

for(i in 1:n){
  for(j in 1:n_stratified){
    tmp_rails_subgroup_ori[[j]] <- c(tmp_rails_subgroup_ori[[j]],
                                     if (is.null(tmp[[i]][[2]][[1]][[j]])) {
                                       NA
                                     } else {
                                       tmp[[i]][[2]][[1]][[j]]
                                     })
    tmp_rails_subgroup_final[[j]] <- c(tmp_rails_subgroup_final[[j]],
                                       if (is.null(tmp[[i]][[3]][[1]][[j]])) {
                                         NA
                                       } else {
                                         tmp[[i]][[3]][[1]][[j]]
                                       })
  }
}

index_nonna <- lapply(tmp_rails_subgroup_ori, function(x){
  if(all(is.na(x))){
    return(FALSE)
  } else {
    return(TRUE)
  }
})

index_sub <- 1:n_stratified
index_sub <- index_sub[unlist(index_nonna)]

# count
count_rails_subgroup_ori <- vector("list",  length(index_sub))
count_rails_subgroup_final <- vector("list",  length(index_sub))

for(j in index_sub){
  tab_ori <- table(unlist(tmp_rails_subgroup_ori[[j]]))
  tab_final <- table(unlist(tmp_rails_subgroup_final[[j]]))
  count_rails_subgroup_ori[[j]] <- append(count_rails_subgroup_ori[[j]],tab_ori)
  count_rails_subgroup_final[[j]] <- append(count_rails_subgroup_final[[j]],tab_final)
}

### Ranking by Subgroups
avg_rank_df_subgroup <- vector("list", n_stratified)
for(j in index_sub){
  rank_models <- vector("list", n)
  for (i in 1:n) {
    rank_models[[i]] <- tmp[[i]][[2]][[1]][[j]]
  }
  all_elements <- unique(unlist(rank_models))
  element_ranks <- setNames(vector("list", length(all_elements)), all_elements)
  # Collect ranks for each element
  for (vec in rank_models) {
    for (i in seq_along(vec)) {
      element <- vec[i]
      element_ranks[[element]] <- c(element_ranks[[element]], i)
    }
  }
  avg_ranks <- sapply(element_ranks, mean)
  avg_ranks <- sort(avg_ranks)
  avg_rank_df_subgroup[[j]] <- data.frame(
    element = names(avg_ranks),
    avg_rank = avg_ranks
  )
}

# Sort the data frame by avg_rank_df_subgroup
for(j in index_sub){
  df2_subgroup_ori <- as.data.frame(as.table(count_rails_subgroup_ori[[j]]), responseName="Freq")
  df2_subgroup_final <- as.data.frame(as.table(count_rails_subgroup_final[[j]]), responseName="Freq")
  
  # Sort the data frame by avg_rank_df_subgroup
  df2_subgroup_ori$Var1 <- factor(df2_subgroup_ori$Var1, levels = avg_rank_df_subgroup[[j]]$element)
  df2_subgroup_final$Var1 <- factor(df2_subgroup_final$Var1, levels = avg_rank_df_subgroup[[j]]$element)
  
  # plot
  p1 <- ggplot(df2_subgroup_ori, aes(x = Var1, y = Freq)) +
    geom_bar(stat = "identity") +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
    )+
    scale_fill_brewer(palette="Dark2")+ 
    scale_x_discrete(limits = avg_rank_df_subgroup[[j]]$element) +
    labs(title=paste0("Subgroup ",names(out_size)[j],", Originally Selected by GVS"), 
         x="Variables", y = "Frequency") 
  
  p2 <- ggplot(df2_subgroup_final, aes(x = Var1, y = Freq)) +
    geom_bar(stat = "identity") +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
    )+
    scale_fill_brewer(palette="Dark2")+ 
    scale_x_discrete(limits = avg_rank_df_subgroup[[j]]$element) +
    labs(title=paste0("Subgroup ",names(out_size)[j],", LIFO after GVS"), 
         x="Variables", y = "Frequency") 
  
  print(p1)
  print(p2)
}

```



